Robust-Adaptive Control of Linear Systems:\\ beyond Quadratic Costs

### 2020-10-06 main.tex
Remove wrong transpose in LMI

\begin{theorem}[Suboptimality bound]
	\label{thm:minimax-regret-bound}
		Under two conditions:
		\begin{enumerate}
			\item a Lipschitz regularity assumption for the reward function $R$:
			\item a stability condition: there exist $P>0,Q_0\in\Real^{p\times p}$, $\rho>0$, and $N_0\in\Natural$ such that
			$$\forall N>N_0,\quad \begin{bmatrix}
			A_N^\transp P + P A_N^\transp + Q_0 & P|D|  \\
			|D|^\transp P & -\rho I_r \\
			\end{bmatrix}< 0;$$
		\end{enumerate}
		we can bound the suboptimality of \Cref{alg:full} with planning budget $K$ as:
		\begin{equation*}
		V(a_\star) - \hat{V}^r(a_K) \leq  \hlr{\underbrace{\Delta_\omega}_{\substack{\text{robustness to}\\ \text{disturbances}}}} + \hlb{\underbrace{\cO\left(\frac{\beta_N(\delta)^2}{\lambda_{\min}(G_{N,\lambda})}\right)}_{\text{estimation error}}} + \hlg{\underbrace{\cO\left(K^{-\frac{\log 1/\gamma}{\log \kappa}}\right)}_{\text{planning error}}} 
		\end{equation*}
		with probability at least $1-\delta$, where $V(a)$ is the optimal expected return after executing an action $a\in\cA$, $a_\star$ is an optimal action, and $\hlr{\Delta_\omega}$ is a constant which corresponds to an irreducible suboptimality suffered from being robust to instantaneous disturbances $\omega(t)$.
\end{theorem}

@@

\begin{corollary}[Asymptotic near-optimality]
	\label{cor:pe}
		Under an additional persistent excitation (PE) assumption
		\begin{align}
		\label{eq:pe}
		\exists \underline{\phi},\overline{\phi}>0: \forall n\geq n_0,\quad \underline{\phi}^2 \leq \lambda_{\min}(\Phi_{n}^\transp\Sigma_{p}^{-1}\Phi_{n}) \leq \overline{\phi}^2,
		\end{align} the stability condition 2. of \Cref{thm:minimax-regret-bound} can be relaxed to its limit: there exist $P,Q_0,\rho$ such that
		$$\begin{bmatrix}
		A(\theta)^\transp P + P A(\theta)^\transp + Q_0 & P|D|  \\
		|D|^\transp P & -\rho I_r \\
		\end{bmatrix}< 0;$$
		and the suboptimality bound takes the more explicit form
		\begin{equation*}
		V(a_\star) - \hat{V}^r(a_K) \leq  \hlr{{\Delta_\omega}} + \hlb{{\cO\left(\frac{\log\left(N^{d/2}/\delta\right)}{N}\right)}} + \hlg{{\cO\left(K^{-\frac{\log 1/\gamma}{\log \kappa}}\right)}} 
		\end{equation*}
		which ensures asymptotic near-optimality when $N\to\infty$ and $K\to\infty$.
\end{corollary}

@@

\begin{lemma}
		If there exist $P>0,Q_0\in\Real^{p\times p}$, $\rho>0$ such that
		\begin{align*}
		\begin{bmatrix}
		A_N^\transp P + P A_N^\transp + Q_0 & P|D|  \\
		|D|^\transp P & -\rho I_r \\
		\end{bmatrix}< 0,
		\end{align*}
		then for all $t> t_N$,
		\[\|\ox(t) - \ux(t)\| \leq \left(C_0 + \cO\left({\frac{\beta_N(\delta)}{\sqrt{\lambda_{\min}(G_{N,\lambda})}}} \right)\right)C_\omega(t), \]
		where $$C_0 = \sqrt{\frac{2\rho\lambda_{\max}(P)}{\lambda_{\min}(P)\lambda_{\min}(Q_0)}},$$ and $$C_\omega(t) = \sup_{\tau\in[0,t]} \|\overline{\omega}(\tau) - \underline{\omega}(\tau)\|_2.$$
\end{lemma}
\begin{proof}
	Let $e = \ox - \ux$. \eqref{eq:interval-predictor} gives the dynamics
	\begin{align*}
	\dot{e} = A_Ne + |\Delta A|(\ox^+ + \ux^-) + |D|(\overline{\omega} - \underline{\omega})
	\end{align*}
	where recall that $|M| = M^+ + M^-$ for any matrix $M\in\Real^{p\times p}$.

	We define the Lyapunov function $V = e^\transp P e$, which is non-negative definite provided that
	$
	P>0,
	$ and compute its derivative
	\begin{align*}
	\dot{V} ={}& X^\transp
	\begin{bmatrix}
	A_N^\transp P + P A_N^\transp + Q & P|D| & P|\Delta A| \\
	|D|^\transp P & -\rho I_r & 0\\
	|\Delta A|^\transp P & 0 & -\alpha I_p
	\end{bmatrix}
	X\\
	& - e^\transp Q e + \alpha |\ux^+ + \ox^-|^2 + \rho |\overline{\omega} - \underline{\omega}|^2
	\end{align*}
	with $X=\begin{bmatrix}
	e & \overline{\omega} - \underline{\omega} &  \ux^+ + \ox^-
	\end{bmatrix}^\transp$, for any $Q\in\Real^{p\times p}$, $\rho,\alpha\in\Real$ .

	Moreover, it holds that $-\ux^+ -\ox^- \leq e \leq \ox^+ + \ux^-$, which implies $|\ux^+ + \ox^-| \leq 2 |e|$. Hence,
	\begin{align*}
	\dot{V} \leq {}& X^\transp
	\underbrace{
		\left[
		\begin{array}{cc|c}
		A_N^\transp P + P A_N^\transp + Q + 4\alpha I_p & P|D| & P|\Delta A| \\
		|D|^\transp P & -\rho I_r & 0\\
		\hline
		|\Delta A|^\transp P & 0 & -\alpha I_p
		\end{array}
		\right]}_{\Upsilon}
	X\\
	& - e^\transp Q e + \rho \|\overline{\omega} - \underline{\omega}\|_2^2
	\end{align*}

@@

	We now examine the condition $\Upsilon \leq 0$.
	We resort to its Schur complement: given $\alpha > 0$, $\Upsilon \leq 0$ if and only if $R \geq S$, where $S= \alpha^{-1}\begin{bmatrix}|\Delta A|^\transp P & 0\end{bmatrix}^\transp \begin{bmatrix}|\Delta A|^\transp P & 0\end{bmatrix}$ and $R$ is the top-left block of $-\Upsilon$:
	\[R = \begin{bmatrix}
	-A_N^\transp P - P A_N^\transp - Q - 4\alpha I_p & -P|D|\\
	-|D|^\transp P & \rho I_r\\
	\end{bmatrix}\]

@@

	Thus, by taking $\alpha = \frac{2\lambda_{\max}(P)^2\|\Delta A\|_F^2}{\lambda_{\min}(Q_0)} = \cO({\frac{\beta_N(\delta)^2}{\lambda_{\min}(G_{N,\lambda})}})$, we can obtain that $S \leq \begin{bmatrix}
	\frac{1}{2}Q_0 & 0\\0 & 0
	\end{bmatrix}$. Thus,
	\[R-S \geq \begin{bmatrix}
	-A_N^\transp P - P A_N^\transp - Q_0 & -P|D|\\
	-|D|^\transp P & \rho I_r\\
	\end{bmatrix} > 0 \]
	as it is assumed in the conditions of the lemma. Hence, under such a choice of $\alpha$ and $Q$, we recover $\Upsilon\leq 0$. \eqref{eq:lyap-bound} follows with $\mu = \frac{\lambda_{\min}(Q)}{\lambda_{\max}(P)} = \frac{\frac{1}{2}\lambda_{\min}(Q_0) - 4\alpha}{\lambda_{\max}(P)}$.
	Finally, we obtain
	\begin{align*}
	\|e(t)\|_2^2 &\leq \lambda_{\min}(P)^{-1} V(t)\\
	& \leq \frac{2\rho\lambda_{\max}(P)/\lambda_{\min}(P)}{\lambda_{\min}(Q_0) - 8\alpha} C_\omega^2(t)\\
	\end{align*}
	Developing at the first order in $\alpha$ gives
	\begin{align*}
	\|e(t)\|_2 &\leq C_0\left(1 + \frac{4\alpha}{\lambda_{\min}(Q_0)} + \cO(\alpha^2)\right)C_\omega(t)\\
	&\leq \left(C_0 + \cO\left({\frac{\beta_N(\delta)^2}{\lambda_{\min}(G_{N,\lambda})}}\right)\right)C_\omega(t)
	\end{align*}
\end{proof}

>>>>

\begin{theorem}[Suboptimality bound]
	\label{thm:minimax-regret-bound}
		Under two conditions:
		\begin{enumerate}
			\item a Lipschitz regularity assumption for the reward function $R$:
			\item a stability condition: there exist $P>0,Q_0\in\Real^{p\times p}$, $\rho>0$, and $N_0\in\Natural$ such that
			$$\forall N>N_0,\quad \begin{bmatrix}
			A_N^\transp P + P A_N + Q_0 & P|D|  \\
			|D|^\transp P & -\rho I_r \\
			\end{bmatrix}< 0;$$
		\end{enumerate}
		we can bound the suboptimality of \Cref{alg:full} with planning budget $K$ as:
		\begin{equation*}
		V(a_\star) - \hat{V}^r(a_K) \leq  \hlr{\underbrace{\Delta_\omega}_{\substack{\text{robustness to}\\ \text{disturbances}}}} + \hlb{\underbrace{\cO\left(\frac{\beta_N(\delta)^2}{\lambda_{\min}(G_{N,\lambda})}\right)}_{\text{estimation error}}} + \hlg{\underbrace{\cO\left(K^{-\frac{\log 1/\gamma}{\log \kappa}}\right)}_{\text{planning error}}} 
		\end{equation*}
		with probability at least $1-\delta$, where $V(a)$ is the optimal expected return after executing an action $a\in\cA$, $a_\star$ is an optimal action, and $\hlr{\Delta_\omega}$ is a constant which corresponds to an irreducible suboptimality suffered from being robust to instantaneous disturbances $\omega(t)$.
\end{theorem}

@@

\begin{corollary}[Asymptotic near-optimality]
	\label{cor:pe}
		Under an additional persistent excitation (PE) assumption
		\begin{align}
		\label{eq:pe}
		\exists \underline{\phi},\overline{\phi}>0: \forall n\geq n_0,\quad \underline{\phi}^2 \leq \lambda_{\min}(\Phi_{n}^\transp\Sigma_{p}^{-1}\Phi_{n}) \leq \overline{\phi}^2,
		\end{align} the stability condition 2. of \Cref{thm:minimax-regret-bound} can be relaxed to its limit: there exist $P,Q_0,\rho$ such that
		$$\begin{bmatrix}
		A(\theta)^\transp P + P A(\theta) + Q_0 & P|D|  \\
		|D|^\transp P & -\rho I_r \\
		\end{bmatrix}< 0;$$
		and the suboptimality bound takes the more explicit form
		\begin{equation*}
		V(a_\star) - \hat{V}^r(a_K) \leq  \hlr{{\Delta_\omega}} + \hlb{{\cO\left(\frac{\log\left(N^{d/2}/\delta\right)}{N}\right)}} + \hlg{{\cO\left(K^{-\frac{\log 1/\gamma}{\log \kappa}}\right)}} 
		\end{equation*}
		which ensures asymptotic near-optimality when $N\to\infty$ and $K\to\infty$.
\end{corollary}

@@

\begin{lemma}
		If there exist $P>0,Q_0\in\Real^{p\times p}$, $\rho>0$ such that
		\begin{align*}
		\begin{bmatrix}
		A_N^\transp P + P A_N + Q_0 & P|D|  \\
		|D|^\transp P & -\rho I_r \\
		\end{bmatrix}< 0,
		\end{align*}
		then for all $t> t_N$,
		\[\|\ox(t) - \ux(t)\| \leq \left(C_0 + \cO\left({\frac{\beta_N(\delta)}{\sqrt{\lambda_{\min}(G_{N,\lambda})}}} \right)\right)C_\omega(t), \]
		where $$C_0 = \sqrt{\frac{2\rho\lambda_{\max}(P)}{\lambda_{\min}(P)\lambda_{\min}(Q_0)}},$$ and $$C_\omega(t) = \sup_{\tau\in[0,t]} \|\overline{\omega}(\tau) - \underline{\omega}(\tau)\|_2.$$
\end{lemma}
\begin{proof}
	Let $e = \ox - \ux$. \eqref{eq:interval-predictor} gives the dynamics
	\begin{align*}
	\dot{e} = A_Ne + |\Delta A|(\ox^+ + \ux^-) + |D|(\overline{\omega} - \underline{\omega})
	\end{align*}
	where recall that $|M| = M^+ + M^-$ for any matrix $M\in\Real^{p\times p}$.

	We define the Lyapunov function $V = e^\transp P e$, which is non-negative definite provided that
	$
	P>0,
	$ and compute its derivative
	\begin{align*}
	\dot{V} ={}& X^\transp
	\begin{bmatrix}
	A_N^\transp P + P A_N + Q & P|D| & P|\Delta A| \\
	|D|^\transp P & -\rho I_r & 0\\
	|\Delta A|^\transp P & 0 & -\alpha I_p
	\end{bmatrix}
	X\\
	& - e^\transp Q e + \alpha |\ux^+ + \ox^-|^2 + \rho |\overline{\omega} - \underline{\omega}|^2
	\end{align*}
	with $X=\begin{bmatrix}
	e & \overline{\omega} - \underline{\omega} &  \ux^+ + \ox^-
	\end{bmatrix}^\transp$, for any $Q\in\Real^{p\times p}$, $\rho,\alpha\in\Real$ .

	Moreover, it holds that $-\ux^+ -\ox^- \leq e \leq \ox^+ + \ux^-$, which implies $|\ux^+ + \ox^-| \leq 2 |e|$. Hence,
	\begin{align*}
	\dot{V} \leq {}& X^\transp
	\underbrace{
		\left[
		\begin{array}{cc|c}
		A_N^\transp P + P A_N + Q + 4\alpha I_p & P|D| & P|\Delta A| \\
		|D|^\transp P & -\rho I_r & 0\\
		\hline
		|\Delta A|^\transp P & 0 & -\alpha I_p
		\end{array}
		\right]}_{\Upsilon}
	X\\
	& - e^\transp Q e + \rho \|\overline{\omega} - \underline{\omega}\|_2^2
	\end{align*}

@@

	We now examine the condition $\Upsilon \leq 0$.
	We resort to its Schur complement: given $\alpha > 0$, $\Upsilon \leq 0$ if and only if $R \geq S$, where $S= \alpha^{-1}\begin{bmatrix}|\Delta A|^\transp P & 0\end{bmatrix}^\transp \begin{bmatrix}|\Delta A|^\transp P & 0\end{bmatrix}$ and $R$ is the top-left block of $-\Upsilon$:
	\[R = \begin{bmatrix}
	-A_N^\transp P - P A_N - Q - 4\alpha I_p & -P|D|\\
	-|D|^\transp P & \rho I_r\\
	\end{bmatrix}\]

@@

	Thus, by taking $\alpha = \frac{2\lambda_{\max}(P)^2\|\Delta A\|_F^2}{\lambda_{\min}(Q_0)} = \cO({\frac{\beta_N(\delta)^2}{\lambda_{\min}(G_{N,\lambda})}})$, we can obtain that $S \leq \begin{bmatrix}
	\frac{1}{2}Q_0 & 0\\0 & 0
	\end{bmatrix}$. Thus,
	\[R-S \geq \begin{bmatrix}
	-A_N^\transp P - P A_N - Q_0 & -P|D|\\
	-|D|^\transp P & \rho I_r\\
	\end{bmatrix} > 0 \]
	as it is assumed in the conditions of the lemma. Hence, under such a choice of $\alpha$ and $Q$, we recover $\Upsilon\leq 0$. \eqref{eq:lyap-bound} follows with $\mu = \frac{\lambda_{\min}(Q)}{\lambda_{\max}(P)} = \frac{\frac{1}{2}\lambda_{\min}(Q_0) - 4\alpha}{\lambda_{\max}(P)}$.
	Finally, we obtain
	\begin{align*}
	\|e(t)\|_2^2 &\leq \lambda_{\min}(P)^{-1} V(t)\\
	& \leq \frac{2\rho\lambda_{\max}(P)/\lambda_{\min}(P)}{\lambda_{\min}(Q_0) - 8\alpha} C_\omega^2(t)\\
	\end{align*}
	Developing at the first order in $\alpha$ gives
	\begin{align*}
	\|e(t)\|_2 &\leq C_0\left(1 + \frac{4\alpha}{\lambda_{\min}(Q_0)} + \cO(\alpha^2)\right)C_\omega(t)\\
	&\leq \left(C_0 + \cO\left({\frac{\beta_N(\delta)^2}{\lambda_{\min}(G_{N,\lambda})}}\right)\right)C_\omega(t)
	\end{align*}
\end{proof}
